---
title: "Genera Hill Numbers"
author: "Alicia M. June"
output: html_document
---
# Introduction

This document contains code for calculating the observed Hill numbers (Fig X, X, X) and the diversity curves (Fig X, X, X). The data is being used is genera ONLY data. 5 samples were used for each member.

Goals:

* Calculate Hill numbers qD for q=0 and q=1
* Create diversity curves for qD for q=0 and q=1
* Extract csvs for diversity curve data, not including extrapolated data

```{r Packages}
library(tidyverse)
library(iNEXT)
library(dplyr)
library(ggplot2)
library(vegan)
```

## Data Preparation
Data is imported as a csv. Samples are in columns, genera in rows. There is also a row for sample ID and the member each sample is from. If a genus was not present in a sample, the abundance count is 0.

```{r Importing data}
genRAW<- read_csv("genRAW.csv", col_names = TRUE)
genRAW
```

### Setting Rows and Columns
```{r Setting Row and Column names}

# made the row name the species name. 
rownames(genRAW)<- genRAW$`Sample ID`
genRAW

#Selecting just the numeric data, converting to data frame
gen_df<- as.data.frame(genRAW[2:141,])
gen_df

#Setting row names to species names again
rownames(gen_df)<- gen_df$`Sample ID`
gen_df


#Just the abundance counts
gen<- (gen_df[,2:16])
gen

rm(genRAW_df)
```


### Formatting for iNEXT
iNEXT works best with lists, or of data frames as a list of matrices. I found that out thanks to this video by
Dr. Óscar M. Chavez, School of Biology, University of Costa Rica:

His video: https://www.youtube.com/watch?v=24PtDhIYg8g

```{r Formatting for iNEXT}
#Step 1: Making each sample into a matrix and also rounding them up

#SM
APSM07MPYB3_rd<- ceiling(as.matrix(as.numeric(gen$`AP SM 07MPYB3`)))
HBSMCG13_rd<- ceiling(as.matrix(as.numeric(gen$`HB SM CG 1:3`)))
APSM07CWEB1_rd<- ceiling(as.matrix(as.numeric(gen$`AP SM 07CWEB1`)))
x230723XBCG_rd<- ceiling(as.matrix(as.numeric(gen$`230723XBCG`)))
HBSMCG31_rd<- ceiling(as.matrix(as.numeric(gen$`HB SM CG 3:1`)))

#RM
NSCG1RM_rd<- ceiling(as.matrix(as.numeric(gen$`NS CG1 RM`)))
x71LW75_rd<- ceiling(as.matrix(as.numeric(gen$`71LW75`)))
NSFW01RM220620_rd<- ceiling(as.matrix(as.numeric(gen$`NSFW-01RM220620`)))
PBDBRM_rd<- ceiling(as.matrix(as.numeric(gen$`PBDB RM`)))
ETRM_rd<- ceiling(as.matrix(as.numeric(gen$`ET RM`)))

#MH
HGMH31MH_rd<- ceiling(as.matrix(as.numeric(gen$`HGMH3-1(MH)`)))
HGMH32_rd<- ceiling(as.matrix(as.numeric(gen$`HGMH3-2`)))
RCMB091723_rd<- ceiling(as.matrix(as.numeric(gen$RCMB091723)))
ETMH_rd<- ceiling(as.matrix(as.numeric(gen$`ET MH`)))
x70LW1RicesPit_rd<- ceiling(as.matrix(as.numeric(gen$`70LW1 (Rice's Pit)`)))

#Step 2: Making the matrices into lists

all_samp_list<- list(APSM07MPYB3= APSM07MPYB3_rd, HBSMCG13= HBSMCG13_rd, APSM07CWEB1= APSM07CWEB1_rd, x230723XBCG= x230723XBCG_rd, HBSMCG31= HBSMCG31_rd, NSCG1RM= NSCG1RM_rd, x71LW75= x71LW75_rd, NSFW01RM220620= NSFW01RM220620_rd, PBDBRM= PBDBRM_rd, ETRM= ETRM_rd, HGMH31MH= HGMH31MH_rd, HGMH32= HGMH32_rd, RCMB091723= RCMB091723_rd,ETMH=ETMH_rd, x70LW1RicesPit= x70LW1RicesPit_rd)

SM_list<- list(APSM07MPYB3= APSM07MPYB3_rd, HBSMCG13= HBSMCG13_rd, APSM07CWEB1= APSM07CWEB1_rd, x230723XBCG= x230723XBCG_rd,  HBSMCG31= HBSMCG31_rd)

RM_list<- list(NSCG1RM= NSCG1RM_rd, x71LW75= x71LW75_rd, NSFW01RM220620= NSFW01RM220620_rd, PBDBRM= PBDBRM_rd,ETRM= ETRM_rd)

MH_list<- list(HGMH31MH= HGMH31MH_rd, HGMH32= HGMH32_rd, RCMB091723= RCMB091723_rd, ETMH=ETMH_rd, x70LW1RicesPit= x70LW1RicesPit_rd)

```


This made it so rarefaction was being calculated as the same intervals that PAST uses. Sage cut off rarefaction at 3000 so I tried to match that. The "size" parameter in the iNEXT object should be "rare_intervals" and "endpoints" and "knots" should both be NULL.

```{r}
rare_intervals<- seq(from =1, to =3000, by= 10)
```


First, calculate the iNEXT object, which includes all order of Hill numbers, including interpolation and
extrapolation.

## Sunken Meadow
```{r SM iNEXT and Data info}
SM_out<- iNEXT(SM_list, q=c(0,1,2), datatype="abundance", size=rare_intervals, endpoint=NULL, knots=NULL, se=TRUE, conf=0.95, nboot=1000)

SM_data_info<- SM_out$DataInfo
SM_data_info
```

**Sunken Meadow: Observed Hill Numbers**
```{r SM Observed Hill Numbers}
SM_samples_observed<- SM_out$AsyEst
write_csv( SM_samples_observed, "SM_gen_Obs.csv")
SM_samples_observed
```

**Sunken Meadow: Rarefaction Data**
```{r SM Rarefaction data }
#Now lets extract our rarefaction data, except extrapolation
SM_rare_data<- SM_out$iNextEst$size_base[SM_out$iNextEst$size_base$Method != "Extrapolation", ]
write_csv(SM_rare_data, "SM_gen_rare.csv")
head(SM_rare_data)
```

## Rushmere
```{r Rushmere iNEXT and Data info}
RM_out<- iNEXT(RM_list, q=c(0,1,2), datatype="abundance", size=rare_intervals, endpoint=NULL, knots=NULL, se=TRUE, conf=0.95, nboot=1000)

RM_data_info<- RM_out$DataInfo
RM_data_info
```

**Rushmere: Observed Hill Numbers**
```{r Rushmere Observed Hill Numbers}

RM_samples_observed<- RM_out$AsyEst
write_csv( RM_samples_observed, "RM_gen_Obs.csv")
RM_samples_observed
```

**Rushmere: Rarefaction Data**
```{r Rushmere Rarefaction Data}
RM_rare_data<- RM_out$iNextEst$size_base[RM_out$iNextEst$size_base$Method != "Extrapolation", ]
write_csv(RM_rare_data, "RM_gen_rare.csv")
head(RM_rare_data)

```

## Moore House
```{r MH iNEXT and Data info}
MH_out<- iNEXT(MH_list, q=c(0,1,2), datatype="abundance", size=rare_intervals, endpoint=NULL, knots=NULL, se=TRUE, conf=0.95, nboot=1000)

MH_data_info<- MH_out$DataInfo
MH_data_info
```

**Moore House: Observed Hill Numbers**
```{r Moore House Observed Hill Numbers}
MH_samples_observed<- MH_out$AsyEst
write_csv( MH_samples_observed, "MH_gen_Obs.csv")
head(MH_samples_observed)
```

**Moore House: Rarefaction Data**
```{r MH Rarefaction Data}
MH_rare_data<- MH_out$iNextEst$size_base[MH_out$iNextEst$size_base$Method != "Extrapolation", ]
write_csv(MH_rare_data, "MH_gen_rare.csv")
head(MH_rare_data)
```

