---
title: "Genera PERMANOVA (all members)"
author: "Alicia June"
output: html_document
---
# Introduction

This document contains code to conduct PERMANOVA comparing the Sunken Meadow, Rushmere, and Moore House members of the Yorktown Formation. The distance matrix is calculated using genera ONLY proportional abundance data. 5 samples were used for each member.

The goals of this code are the following:

* Run PERMANOVA on genera ONLY data
* Calculate centroids using DCA coordinates
* Calculate distance matrix with using proportional abundance data to account for sample size
* 5 trials of distance matrix
* 10 trials of PERMANOVA, permutations = 10000

```{r Packages}
library(tidyverse)
library(iNEXT)
library(dplyr)
library(ggplot2)
library(vegan)
```

## Data Preparation: Abundance Data
Abundance data is imported as a csv. Samples are in rows, genera in columns. There is also a column for sample ID and the member each sample is from. If a genus was not present in a sample, the abundance count is 0.
```{r Data prep}
gen_raw<- read_csv("gen_tran.csv", col_names = TRUE)
gen_raw
```

### Just Numeric Data
```{r Just nums}
gen_num<- gen_raw[,3:142]
gen_num
```

### Getting Proportions
```{r}
#Converting to proportions
perm_data<- (100*(gen_num/rowSums(gen_num)))
perm_data
```

## Data Preparation: DCA Data
```{r Import DCA data}
dcadata<- read_csv("gen_dca.csv", col_names = TRUE)
dcadata

#Extracting DCA scores
dca_scores<- dcadata[,3:4]
dca_scores
```

## Plotting

Step 1: Find centroids
```{r Finding Centroids}
dcagroup_centroids <- data.frame(
  Member = c("SM", "RM", "MH"),
  Centroid_X = c(mean(dca_scores$`Axis 1`[dcadata$Member == "SM"]),
                 mean(dca_scores$`Axis 1`[dcadata$Member == "RM"]),
                 mean(dca_scores$`Axis 1`[dcadata$Member == "MH"])),
  Centroid_Y = c(mean(dca_scores$`Axis 2`[dcadata$Member == "SM"]),
                 mean(dca_scores$`Axis 2`[dcadata$Member == "RM"]),
                 mean(dca_scores$`Axis 2`[dcadata$Member == "MH"]))
)

```


 Step 2: Create df for ggplot
```{r Create df for ggplot}
dcaplot_data <-data.frame(
  Member = dcadata$Member,
  DC1 = dca_scores$`Axis 1`,
  DC2 = dca_scores$`Axis 2`,
  xend = c(rep(dcagroup_centroids[1,2],5), rep(dcagroup_centroids[2,2],5), rep(dcagroup_centroids[3,2],5)),
  yend = c(rep(dcagroup_centroids[1,3],5), rep(dcagroup_centroids[2,3],5), rep(dcagroup_centroids[3,3],5))
)

dcaplot_data
```

Step 3: Plot the DCA

* Plot samples, coloring based on location
* Draw the ellipse and the segments

```{r DCA plot}
ggplot(dcaplot_data, aes(DC1,DC2))+
  geom_point(aes(color = Member), size = 2) +
  stat_ellipse(geom = "polygon", alpha = 0.04, aes(group = Member),
               color = "black", fill = "blue")+
  geom_point(data = dcagroup_centroids, aes(x = Centroid_X, y = Centroid_Y),
             color = "black", size = 2, shape = 7)+
  geom_segment(data = dcaplot_data, aes(x =DC1, y = DC2, 
                                        xend = xend, yend = yend, color = Member), alpha = 0.5)

```

# Running PERMANOVA

## Distance Matrix
```{r}
perm_dist<-vegdist(perm_data, method = 'bray')
perm_dist

```

Creating df of distance matrix
```{r}
#Create empty matrix
dist_mat= matrix(0, nrow = 15, ncol = 15)

#Setting below diagonal to perm_dost values, and above the diagonal to empty
dist_mat[lower.tri(dist_mat)] <- perm_dist
dist_mat[upper.tri(dist_mat)]<- ""
dist_mat

#making it a df
perm_dist_df<- as.data.frame(dist_mat, row.names = NULL)
perm_dist_df

#setting col and rownames so we can see what distance we are calculating
rownames(perm_dist_df)<- gen_raw$`Sample ID`
colnames(perm_dist_df)<- gen_raw$`Sample ID`

write_csv(perm_dist_df, "gen_dist_all.csv")
perm_dist_df
```

## Assumption
```{r}
#Assumptions- I don't think this applies to us, but keeping it in.
dispersion<-betadisper(perm_dist, group=gen_raw$Member,type = "centroid")

plot(dispersion)

anova(dispersion)

```

## PERMANOVA Trials

Repeated trials of PERMANOVA at 10,000 permutations. Will average p-values later. 

Trial 00
```{r Trial 00}
Perma_result <- adonis2(perm_dist~as.factor(dcaplot_data$Member), data = perm_dist,
                        permutations = 10000)
Perma_result

```

Trial 01
```{r Trial 01}
Perma_result1 <- adonis2(perm_dist~as.factor(dcaplot_data$Member), data = perm_dist,
                         permutations = 10000)
Perma_result1

#Extracting perm results
permustats_results1<- permustats(Perma_result1)
#permustats_results1

#calculating mean and se
means_and_se1<- mean_se(permustats_results1$permutations, mult = 1)

```
Trial 02
```{r Trial 02}
Perma_result2 <- adonis2(perm_dist~as.factor(dcaplot_data$Member), data = perm_dist,
                         permutations = 10000)
Perma_result2

#Extracting perm results
permustats_results2<- permustats(Perma_result2)
#permustats_results2

#calculating mean and se
means_and_se2<- mean_se(permustats_results2$permutations, mult = 1)

```


Trial 03
```{r Trial 03}
Perma_result3 <- adonis2(perm_dist~as.factor(dcaplot_data$Member), data = perm_dist,
                         permutations = 10000)
Perma_result3

#Extracting perm results
permustats_results3<- permustats(Perma_result3)
#permustats_results3

#calculating mean and se
means_and_se3<- mean_se(permustats_results3$permutations, mult = 1)
```


Trial 04
```{r Trial 04}
Perma_result4 <- adonis2(perm_dist~as.factor(dcaplot_data$Member), data = perm_dist,
                         permutations = 10000)
Perma_result4

#Extracting perm results
permustats_results4<- permustats(Perma_result4)
#permustats_results4

#calculating mean and se
means_and_se4<- mean_se(permustats_results4$permutations, mult = 1)
```

Trial 05
```{r Trial 05}
Perma_result5 <- adonis2(perm_dist~as.factor(dcaplot_data$Member), data = perm_dist,
                         permutations = 10000)
Perma_result5

#Extracting perm results
permustats_results5<- permustats(Perma_result5)
#permustats_results5

#calculating mean and se
means_and_se5<- mean_se(permustats_results5$permutations, mult = 1)
```

Trial 06
```{r Trial 06}
Perma_result6 <- adonis2(perm_dist~as.factor(dcaplot_data$Member), data = perm_dist,
                         permutations = 10000)
Perma_result6

#Extracting perm results
permustats_results6<- permustats(Perma_result6)
#permustats_results6

#calculating mean and se
means_and_se6<- mean_se(permustats_results6$permutations, mult = 1)
```

Trial 07
```{r Trial 07}
Perma_result7 <- adonis2(perm_dist~as.factor(dcaplot_data$Member), data = perm_dist,
                         permutations = 10000)
Perma_result7

#Extracting perm results
permustats_results7<- permustats(Perma_result7)
#permustats_results7

#calculating mean and se
means_and_se7<- mean_se(permustats_results7$permutations, mult = 1)
```

Trial 08
```{r Trial 08}
Perma_result8 <- adonis2(perm_dist~as.factor(dcaplot_data$Member), data = perm_dist,
                         permutations = 10000)
Perma_result8

#Extracting perm results
permustats_results8<- permustats(Perma_result8)
#permustats_results8

#calculating mean and se
means_and_se8<- mean_se(permustats_results8$permutations, mult = 1)
```

Trial 09
```{r Trial 09}
Perma_result9 <- adonis2(perm_dist~as.factor(dcaplot_data$Member), data = perm_dist,
                         permutations = 10000)
Perma_result9

#Extracting perm results
permustats_results9<- permustats(Perma_result9)
#permustats_results9

#calculating mean and se
means_and_se9<- mean_se(permustats_results9$permutations, mult = 1)
```

Trial 10
```{r Trial 10}
Perma_result10 <- adonis2(perm_dist~as.factor(dcaplot_data$Member), data = perm_dist,
                         permutations = 10000)
Perma_result10

#Extracting perm results
permustats_results10<- permustats(Perma_result10)
#permustats_results10

#calculating mean and se
means_and_se10<- mean_se(permustats_results10$permutations, mult = 1)
```





