---
title: "spONLY V3 Hill Rare 03_17_2025"
author: "Alicia June"
date: "2025-03-17"
output: html_document
---

# Introduction

This document is V3 of conducting rarefaction using Hill Numbers. The data is being used is species ONLY data (03-17-2025). V3 uses 5 samples for each member. It excludes the following samples: 

* AP SM 07MPEA2 (Sunken Meadow)
* VMNH RM 1 (Rushmere)
* VMNH MH1 (Moore House)

Goals:

* Create Hill Rarefaction curves for the order q=0 and q=1
* Extract csvs for rarefaction data, not including rarefaction

```{r Packages}
library(tidyverse)
library(iNEXT)
library(dplyr)
library(ggplot2)
library(vegan)
```

## Data Preparation
Data is imported as a csv. Samples are in columns, species in rows. There is also a row for sample ID and the member each sample is from. If a species was not present in a sample, the abundance count is 0.

```{r Importing data}
spONLYraw<- read_csv("spONLY031325.csv", col_names = TRUE)
spONLYraw
```
### Setting Rows and Columns
```{r Setting Row and Column names}

# made the row name the species name. 
rownames(spONLYraw)<- spONLYraw$`Sample ID`
spONLYraw

#Selecting just the numeric data, converting to data frame
spONLY_df<- as.data.frame(spONLYraw[2:197,])
spONLY_df

#Setting row names to species names again
rownames(spONLY_df)<- spONLY_df$`Sample ID`
spONLY_df


#Just the abundance counts
spONLY<- (spONLY_df[,2:16])
spONLY
```

### Formatting for iNEXT
iNEXT works best with lists, or of data frames as a list of matrices. I found that out thanks to this video by
Dr. Óscar M. Chavez, School of Biology, University of Costa Rica:

His video: https://www.youtube.com/watch?v=24PtDhIYg8g

```{r Formatting for iNEXT}
#Step 1: Making each sample into a matrix and also rounding them up

#SM
APSM07MPYB3_rd<- ceiling(as.matrix(as.numeric(spONLY$`AP SM 07MPYB3`)))
HBSMCG13_rd<- ceiling(as.matrix(as.numeric(spONLY$`HB SM CG 1:3`)))
APSM07CWEB1_rd<- ceiling(as.matrix(as.numeric(spONLY$`AP SM 07CWEB1`)))
x230723XBCG_rd<- ceiling(as.matrix(as.numeric(spONLY$`230723XBCG`)))
HBSMCG31_rd<- ceiling(as.matrix(as.numeric(spONLY$`HB SM CG 3:1`)))

#RM
NSCG1RM_rd<- ceiling(as.matrix(as.numeric(spONLY$`NS CG1 RM`)))
x71LW75_rd<- ceiling(as.matrix(as.numeric(spONLY$`71LW75`)))
NSFW01RM220620_rd<- ceiling(as.matrix(as.numeric(spONLY$`NSFW-01RM220620`)))
PBDBRM_rd<- ceiling(as.matrix(as.numeric(spONLY$`PBDB RM`)))
ETRM_rd<- ceiling(as.matrix(as.numeric(spONLY$`ET RM`)))

#MH
HGMH31MH_rd<- ceiling(as.matrix(as.numeric(spONLY$`HGMH3-1(MH)`)))
HGMH32_rd<- ceiling(as.matrix(as.numeric(spONLY$`HGMH3-2`)))
RCMB091723_rd<- ceiling(as.matrix(as.numeric(spONLY$RCMB091723)))
ETMH_rd<- ceiling(as.matrix(as.numeric(spONLY$`ET MH`)))
x70LW1RicesPit_rd<- ceiling(as.matrix(as.numeric(spONLY$`70LW1 (Rice's Pit)`)))

#Step 2: Making the matrices into lists

all_samp_list<- list(APSM07MPYB3= APSM07MPYB3_rd, HBSMCG13= HBSMCG13_rd, APSM07CWEB1= APSM07CWEB1_rd, x230723XBCG= x230723XBCG_rd, HBSMCG31= HBSMCG31_rd, NSCG1RM= NSCG1RM_rd, x71LW75= x71LW75_rd, NSFW01RM220620= NSFW01RM220620_rd, PBDBRM= PBDBRM_rd, ETRM= ETRM_rd, HGMH31MH= HGMH31MH_rd, HGMH32= HGMH32_rd, RCMB091723= RCMB091723_rd,ETMH=ETMH_rd, x70LW1RicesPit= x70LW1RicesPit_rd)

SM_list<- list(APSM07MPYB3= APSM07MPYB3_rd, HBSMCG13= HBSMCG13_rd, APSM07CWEB1= APSM07CWEB1_rd, x230723XBCG= x230723XBCG_rd,  HBSMCG31= HBSMCG31_rd)

RM_list<- list(NSCG1RM= NSCG1RM_rd, x71LW75= x71LW75_rd, NSFW01RM220620= NSFW01RM220620_rd, PBDBRM= PBDBRM_rd,ETRM= ETRM_rd)

MH_list<- list(HGMH31MH= HGMH31MH_rd, HGMH32= HGMH32_rd, RCMB091723= RCMB091723_rd, ETMH=ETMH_rd, x70LW1RicesPit= x70LW1RicesPit_rd)

```

This made it so rarefaction was being calculated as the same intervals that PAST uses. Sage cut off rarefaction at 3000 so I tried to match that. The "size" parameter in the iNEXT object should be "rare_intervals" and "endpoints" and "knots" should both be NULL.

```{r}
rare_intervals<- seq(from =1, to =3000, by= 10)
```

First, calculate the iNEXT object, which includes all order of Hill numbers, including interpolation and
extrapolation.

## Sunken Meadow
```{r SM iNEXT and Data info}
SM_out<- iNEXT(SM_list, q=c(0,1,2), datatype="abundance", size=rare_intervals, endpoint=NULL, knots=NULL, se=TRUE, conf=0.95, nboot=1000)

SM_data_info<- SM_out$DataInfo
SM_data_info
```

**Sunken Meadow: Observed Hill Numbers**
```{r SM Observed Hill Numbers}
SM_samples_observed<- SM_out$AsyEst
write_csv( SM_samples_observed, "SM_spONLY_ObsV3.csv")
SM_samples_observed
```

**Sunken Meadow: Rarefaction Data**
```{r SM Rarefaction data }
#Now lets extract our rarefaction data, except extrapolation
SM_rare_data<- SM_out$iNextEst$size_base[SM_out$iNextEst$size_base$Method != "Extrapolation", ]
write_csv(SM_rare_data, "SM_spONLY_rareV3.csv")
head(SM_rare_data)
```

## Rushmere
```{r Rushmere iNEXT and Data info}
RM_out<- iNEXT(RM_list, q=c(0,1,2), datatype="abundance", size=rare_intervals, endpoint=NULL, knots=NULL, se=TRUE, conf=0.95, nboot=1000)

RM_data_info<- RM_out$DataInfo
RM_data_info
```

**Rushmere: Observed Hill Numbers**
```{r Rushmere Observed Hill Numbers}

RM_samples_observed<- RM_out$AsyEst
write_csv( RM_samples_observed, "RM_spONLY_ObsV3.csv")
RM_samples_observed
```

**Rushmere: Rarefaction Data**
```{r Rushmere Rarefaction Data}
RM_rare_data<- RM_out$iNextEst$size_base[RM_out$iNextEst$size_base$Method != "Extrapolation", ]
write_csv(RM_rare_data, "RM_spONLY_rareV3.csv")
head(RM_rare_data)

```

## Moore House
```{r MH iNEXT and Data info}
MH_out<- iNEXT(MH_list, q=c(0,1,2), datatype="abundance", size=rare_intervals, endpoint=NULL, knots=NULL, se=TRUE, conf=0.95, nboot=1000)

MH_data_info<- MH_out$DataInfo
MH_data_info
```

**Moore House: Observed Hill Numbers**
```{r Moore House Observed Hill Numbers}
MH_samples_observed<- MH_out$AsyEst
write_csv( MH_samples_observed, "MH_spONLY_ObsV3.csv")
head(MH_samples_observed)
```

**Moore House: Rarefaction Data**
```{r MH Rarefaction Data}
MH_rare_data<- MH_out$iNextEst$size_base[MH_out$iNextEst$size_base$Method != "Extrapolation", ]
write_csv(MH_rare_data, "MH_spONLY_rareV3.csv")
head(MH_rare_data)
```




